---
title: "Data Exploration"
format: html
editor: visual
execute:
  echo: false
  warning: false
  message: false
  error: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(readxl)
library(gtsummary)
library(corrplot)
library(patchwork)
```

## Read Data

```{r}
#| label: read-data

# Clinical has a 2-row header: row 1 = section group labels (merged cells),
# row 2 = actual column names. Extract row 2 as col names, then read from row 3.
clinical_col_names <- read_excel(
  "raw-data/Clinical_and_Other_Features.xlsx",
  skip = 1, n_max = 1, col_names = FALSE
) |>
  unlist(use.names = FALSE)

clinical <- read_excel(
  "raw-data/Clinical_and_Other_Features.xlsx",
  skip = 2,
  col_names = clinical_col_names
)
 
clin <- clinical %>% janitor::clean_names()
clin <- clin[2:nrow(clin), ]

imaging <- read_excel("raw-data/Imaging_Features.xlsx")
```

# Outcomes

```{r}
clin <- clin %>%
  mutate(
    response = factor(
      pathologic_response_to_neoadjuvant_therapy,
      levels = 1:5,
      labels = c(
        "Complete response",
        "Not complete response",
        "DCIS only remaining",
        "LCIS only remaining",
        "Treatment response assessment unavailable"
      )
    ),
    response_binary = factor(
      case_when(
        response %in% c("Complete response", "DCIS only remaining", "LCIS only remaining") ~ "Response",
        response == "Not complete response" ~ "No response",
        .default = NA
      ),
      levels = c("Response", "No response")
    )
  )
```


```{r}
clin %>% select(response, response_binary) %>%
  tbl_summary()
```

```{r}
#| label: subtype

clin <- clin %>%
  mutate(
    across(c(er, pr, her2), as.numeric),
    subtype = factor(
      case_when(
        (er == 1 | pr == 1) & her2 == 0 ~ "Luminal A",
        (er == 1 | pr == 1) & her2 == 1 ~ "Luminal B",
        er == 0 & pr == 0 & her2 == 1   ~ "HER2",
        er == 0 & pr == 0 & her2 == 0   ~ "Triple-negative"
      ),
      levels = c("Luminal A", "Luminal B", "HER2", "Triple-negative")
    ),
    subtype_luminal_a = factor(
      case_when(
        subtype == "Luminal A" ~ "Luminal A",
        !is.na(subtype)        ~ "Other"
      ),
      levels = c("Luminal A", "Other")
    )
  )
```

```{r}
clin %>% select(subtype) %>%
  tbl_summary()
```



# Imaging Features


```{r}
names(imaging)
``` 

```{r}
#| label: imaging-corr

df_imaging <- imaging |>
  select(where(is.numeric)) |>
  janitor::clean_names()

colnames(df_imaging) <- abbreviate(
  colnames(df_imaging),
  minlength = 3,
  use.classes = TRUE
)

cor_matrix <- cor(df_imaging, use = "pairwise.complete.obs")

corrplot(
  cor_matrix,
  method   = "color",
  type     = "upper",
  order    = "hclust",
  tl.cex   = 0.6,
  tl.col   = "black",
  addCoef.col = NULL,
  col      = colorRampPalette(c("#2166AC", "white", "#D6604D"))(200),
  title    = "Imaging Feature Correlations",
  mar      = c(0, 0, 2, 0)
)
```





## By Group

Group descriptions:

| Group | n | What it is |
|---|---|---|
| Kinetic_DT | 18 | F1_DT delay-time curve features (both sequences, all T11/T12 thresholds) |
| Tumor_Morphology | 10 | Size/shape + BE descriptors (major axis, volume, ASD, MF, solidity, elongation, Euler) |
| GLCM_Tumor | 22 | Haralick texture on tumor |
| Kinetic_Curve | 4 | Characteristic-curve kinetics (max enhancement, time-to-peak, uptake, washout) |
| FGT_Volume | 49 | Breast/tissue volume, density, enhancement-ratio features (T1 + PostCon) |
| Enh_Variance | 4 | Tumor enhancement-variance dynamics |
| Kinetic_Grouping | 76 | All Grouping_based_* slope features (tumor 2D/3D, tissue T1/PostCon) |
| Spatial | 4 | Moran's I + enhancement-cluster features |
| DFT_Moment | 22 | DFT moment invariants (2D/3D) |
| Dynamic_Texture | 22 | DHOG/DLBP features |
| GLCM_Tissue_T1 | 22 | Haralick texture on tissue (T1) |
| GLCM_Tissue_PostCon | 22 | Haralick texture on tissue (PostCon) |
| Margin_Uptake | 5 | Margin gradient, uptake variance, RGH variance |
| SER_Tumor | 36 | Parametric-map SER features on tumor |
| PE_Tumor | 27 | Parametric-map PE features on tumor |
| Washin_Tumor | 26 | Parametric-map Washin features on tumor |
| SER_Tissue_T1 | 27 | SER features on tissue (T1) |
| PE_Tissue_T1 | 27 | PE features on tissue (T1) |
| Washin_Tissue_T1 | 26 | Washin features on tissue (T1) |
| SER_Tissue_PostCon | 27 | SER features on tissue (PostCon) |
| PE_Tissue_PostCon | 27 | PE features on tissue (PostCon) |
| Washin_Tissue_PostCon | 26 | Washin features on tissue (PostCon) |

```{r}
#| label: group-setup

gr <- read.csv(here::here("raw-data", "variable_groups.csv"))

# clean variable names to match df_imaging
gr <- gr |>
  mutate(variable_clean = janitor::make_clean_names(variable))
```

```{r}
#| label: intra-group-corr
#| fig-height: 8

# Full correlation matrix (unabbreviated names for joining)
df_imaging_raw <- imaging |>
  select(where(is.numeric)) |>
  janitor::clean_names()

cor_full <- cor(df_imaging_raw, use = "pairwise.complete.obs")

groups <- unique(gr$group)
groups <- groups[groups != "ID"]  # drop patient ID

for (grp in groups) {
  vars <- gr$variable_clean[gr$group == grp]
  vars <- intersect(vars, colnames(df_imaging_raw))
  if (length(vars) < 2) next

  mat <- cor_full[vars, vars]

  corrplot(
    mat,
    method  = "color",
    type    = "upper",
    order   = "hclust",
    tl.cex  = 0.55,
    tl.col  = "black",
    col     = colorRampPalette(c("#2166AC", "white", "#D6604D"))(200),
    title   = paste("Intra-group:", grp),
    mar     = c(0, 0, 2, 0)
  )
}
```

```{r}
#| label: inter-group-corr
#| fig-height: 6
#| fig-width: 12

group_labels <- groups

# Inter-group block matrix (mean |r| between each pair of groups)
block_cor <- matrix(NA, length(group_labels), length(group_labels),
                    dimnames = list(group_labels, group_labels))

for (i in seq_along(group_labels)) {
  for (j in seq_along(group_labels)) {
    vi <- intersect(gr$variable_clean[gr$group == group_labels[i]], colnames(df_imaging_raw))
    vj <- intersect(gr$variable_clean[gr$group == group_labels[j]], colnames(df_imaging_raw))
    if (length(vi) == 0 || length(vj) == 0) next
    block_cor[i, j] <- mean(abs(cor_full[vi, vj]), na.rm = TRUE)
  }
}

# Intra-group mean |r| (upper triangle only, no self-correlation)
intra <- tibble(
  group = group_labels,
  mean_abs_r = map_dbl(group_labels, function(g) {
    v <- intersect(gr$variable_clean[gr$group == g], colnames(df_imaging_raw))
    if (length(v) < 2) return(NA_real_)
    m <- cor_full[v, v]
    mean(abs(m[upper.tri(m)]), na.rm = TRUE)
  })
) |>
  filter(!is.na(mean_abs_r)) |>
  arrange(desc(mean_abs_r))

# Bar chart: intra-group
p_bar <- ggplot(intra, aes(x = reorder(group, mean_abs_r), y = mean_abs_r)) +
  geom_col(fill = "#2166AC", width = 0.6) +
  geom_text(aes(label = round(mean_abs_r, 2)), hjust = -0.15, size = 3) +
  coord_flip(ylim = c(0, 1)) +
  labs(
    title = "Intra-group mean |r|",
    x = NULL, y = "Mean |r|"
  ) +
  theme_minimal(base_size = 11)

# Heatmap: inter-group as ggplot tile (upper triangle only)
inter_df <- as.data.frame(as.table(block_cor)) |>
  rename(group_x = Var1, group_y = Var2, mean_abs_r = Freq) |>
  filter(
    match(as.character(group_x), group_labels) <= match(as.character(group_y), group_labels)
  ) |>
  mutate(
    group_x = factor(group_x, levels = group_labels),
    group_y = factor(group_y, levels = rev(group_labels))
  )

p_heat <- ggplot(inter_df, aes(x = group_x, y = group_y, fill = mean_abs_r)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(mean_abs_r, 2)), size = 2.8) +
  scale_fill_gradient(low = "white", high = "#D6604D", limits = c(0, 1), na.value = "grey95") +
  labs(title = "Inter-group mean |r|", x = NULL, y = NULL, fill = "Mean |r|") +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_bar + p_heat + plot_layout(widths = c(1, 1.4))
```


# Association with Outcome


```{r}

```